\\textbf{\\textsc{Supplementary Material} \\linebreak \\linebreak Not just for programmers: How GitHub can accelerate collaborative and reproducible research in ecology and evolution \\linebreak \\linebreak Methods to create \\textsl{the scatterblob} figure}

### 2023-02-23 R/Supp_Info_1.Rmd
streamline Appendix S1 and S2 and code chunks (somewhat)
This commit edits S1 to be more concise, and moves some content from S2 into S1. This was done so the descriptions about how each index was computed is in one section. The Appendix S2 is now mainly the instructions to treat the data and plot it to generate Figure 2 from the main text. This code has also been edited down slightly, to only show the code needed to generate final Figure 2, without exercises in colour choices.

---
output:
  bookdown::pdf_document2:
      latex_engine: xelatex
      number_sections: false
      toc: false
mainfont: "Times New Roman"
geometry: margin = 1in
spacing: double
fontsize: 11pt
editor_options:
  markdown:
    wrap: sentence
csl: https://raw.githubusercontent.com/marlonecobos/kuenm/master/ecography.csl
link-citations: yes
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}\doublespacing
   - \usepackage{parskip}
   - \setlength{\parindent}{4em}
   - \setlength{\parskip}{0em}
   - \usepackage{caption}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

@@

We aimed at distributing the use cases discussed in our study across gradients of technical difficulty and collaboration.
For this, we [...].

We then [\...].

We used Google Docs and Google Sheets to manipulate data, and plotted results using `R` version 4.02.

\noindent

@@

**Goal**: This is a summary figure, e.g., here are all the ways and here are the entry points with low technical difficulty, some for beginners and some more advanced ways to use GitHub.

**Highlight**: There are ways to use github that allow different degrees of collaboration that do not require too many technical skills.

## The data

Data was assembled collaboratively using Google docs, then manually saved as a "csv" file called `scatterblob_data_raw.csv` in the `data` folder.

The collaborative [google spread sheet](https://docs.google.com/spreadsheets/d/1zCRgctjfjE2xZN9MkqfFh222K6bvVvA3gT0UdkL_CS8/edit#gid=0), features a first column listing the ways to use GitHub that were addressed in the manuscript (`r paste0(uses[1:13], collapse = "; " )`).

The column *`r column_names[6]`* corresponds to the degree of difficulty that was perceived by the authors when assembling the data, based on personal experience.
We quickly realized this varied greatly across authors, and that this variation seemed to depend on which github tools were previously known or mastered by the authors.

The columns tagged as *Cognitive load -* showcase a different way to assess technical difficulty of GitHub uses, measured as cognitive load (explained in the next section).

The columns tagged as *degree of collaboration* showcase the span of collaboration that can be achieved with each tool, as perceived by the authors.

### Measuring degree of technical difficulty

[Cognitive load](https://docs.google.com/spreadsheets/d/1YipCCdg5Z8w-IQO67rFfMBSTpLjO-_7HiW44gqzTXPM/edit#gid=0) was measured as follows:

## Reading the data CSV file:

```{r}
scatterblob_data <- read_csv(here("data", "scatterblob_data_raw.csv"))
head(scatterblob_data)
#  View(scatterblob_data)
```

Cleaning the raw data and saving it as a `my_data` object:

```{r}
names(scatterblob_data)
scatterblob_data$x1 <- scatterblob_data$`Min degree of collaboration X1`
scatterblob_data$x2 <- scatterblob_data$`Max degree of collaboration X2`
scatterblob_data$y_real <- scatterblob_data$`Cognitive load - real - non-overlapping`
scatterblob_data$y_seq <- scatterblob_data$`Cognitive load - sequential`
my_data <- scatterblob_data[1:13,c("Ways to use GitHub","y_seq", "y_real", "x1", "x2")]
my_data
```

## Plotting the data

@@

### 1) Using raw cognitive values.

Define colors:

```{r}
some_colors <- viridis::turbo(n = nrow(my_data), alpha = 0.9)

some_colors1 <- viridis::turbo(n = 30, alpha = 0.9)

some_colors2 <- viridis::magma(n = 30, alpha = 0.9)

plot(1:30, 1:30, col = some_colors2, pch = 8)
```

```{r}
my_colors <- some_colors1[c(2:3, 5,
                           8, 10, 12,
                           20, 21, 22, 23, 24,
                           28, 30)]

plot(1:length(my_colors), 
     1:length(my_colors), 
     col = my_colors, pch = 8)

```

```{r}
###############################################################
# start the pdf/png connection and set graphical parameters
###############################################################
# plot_format <- "pdf"
# plot_format <- "svg"
 plot_format <- "png"

if (plot_format == "pdf") {
  pdf(file = here(outpath, "scatterblob_0.pdf"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "svg") {
  svg(filename = here(outpath, "scatterblob_0.svg"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "png") {
  png(file = here(outpath, "scatterblob_0.png"),
      width = 8.25,
      height = 6,
      units = "in",
      res = 300,
      bg = "white")  
}
par(xpd = NA,
    mai = c(1.02, 1, 0.82, 4),
    bty="n")
###############################################################
# create the plot background
###############################################################
plot(x = c(0.5, 5.5),
     y = c(1.75, 11.5),
     xlab = "",
     ylab = "",
     col = "white",
     # "hide" the tick labels so we can put some words instead of numbers:
     col.axis = "white",
     tck = 0.02)
title(ylab="Technical difficulty", line = 4, cex.lab = 1)
title(xlab="Degree of collaboration", line = 3.5, cex.lab = 1)
###############################################################
# add tick labels
###############################################################
text(x = 1:5,
     y = c(1, rep(1.05, 4)),
     cex = 0.7,
     labels = c("Personal",
                "Lab \nMembers",
                "Other \nLabs",
                "EEB \nCommunity",
                "All \nUsers"))
text(x = c(1, 3, 5),
     y = 0.5,
     labels = c("Low", "Medium", "High"))
# Technical difficulty levels from https://www.londonschool.com/level-scale/
text(x = 0.2,
     y = c(2, 4, 6, 8, 10),
     cex = 0.7,
     adj = 1,
     labels = c("Elementary",
                "Low \nIntermediate",
                "Intermediate",
                "Pre \nAdvanced",
                "Very \nAdvanced"))
###############################################################
# plot scatter blobs
###############################################################
for (i in seq(nrow(my_data))) {
  X0 <- my_data$x1[i]
  X1 <- my_data$x2[i]
  Y <- my_data$y_real[i]
  if (X0 == X1) {
    X0 <- X0 - 0.01
    X1 <- X1 + 0.01
   }
  segments(x0 = X0,
           x1 = X1,
           y0 = Y,
           col = my_colors[i],
           lwd = 13)
}
###############################################################
# add github uses text
###############################################################
text(x = 5.5,
    y = my_data$y_real,
    labels = my_data$`Ways to use GitHub`,
    adj = 0,
    col = my_colors)
dev.off()
```

![Degree of collaboration VS technical difficulty (real, non-overlapping)](../content/images/scatterblob_0.png)

### 2) Distributing cognitive load evenly:

We tried different color palettes from two different R packages: [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html#the-color-scales) and [wesanderson](https://github.com/karthik/wesanderson#readme):

```{r}
some_colors3 <- viridis::turbo(n = 100, alpha = 0.9)
some_colors4 <- wesanderson::wes_palette("Zissou1", 75, type = "continuous")
some_colors5 <- wesanderson::wes_palette("Darjeeling1", 75, type = "continuous")
some_colors6 <- wesanderson::wes_palette("Darjeeling2", 75, type = "continuous")
some_colors7 <- wesanderson::wes_palette("Royal2", 75, type = "continuous")
plot(1:75, 1:75, col = some_colors7, pch = 8)
```

We agreed that viridis-turbo was the nicest palette for the purposes of our figure.

```{r}
my_colors <- some_colors3[c(2, 5, 15,
                           26, 28, 30,
                           68, 72, 76, 80, 84,
                           94, 100)]
plot_name <- "scatterblob_1-viridis-turbo"
# my_colors <- some_colors4[round(my_data$y_real*10)-20]
# plot_name <- "scatterblob_1-zissou1"
# my_colors <- some_colors5[round(my_data$y_real*10)-20]
# plot_name <- "scatterblob_1-darjeeling1"
# my_colors <- some_colors6[round(my_data$y_real*10)-20]
# plot_name <- "scatterblob_1-darjeeling2"
# my_colors <- some_colors7[round(my_data$y_real*10)-20]
# plot_name <- "scatterblob_1-royal2"
###############################################################
# start the pdf/png/svg connection and set graphical parameters
###############################################################
# plot_format <- "pdf"
plot_format <- "png"
# plot_format <- "svg"

if (plot_format == "pdf") {
  pdf(file = here(outpath, paste0(plot_name, ".pdf")),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "svg") {
  svg(filename = here(outpath, paste0(plot_name, ".svg")),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "png") {
  png(file = here(outpath, paste0(plot_name, ".png")),
      width = 8.25,
      height = 6,
      units = "in",
      res = 300,
      bg = "white")  
}
par(xpd = NA,
    mai = c(1.02, 1, 0.82, 4),
    bty="n")
###############################################################
# create the plot background
###############################################################
plot(x = c(0.5, 5.5),
     y = c(0.5, 13.5),
     xlab = "",
     ylab = "",
     col = "white",
     # "hide" the tick labels so we can put some words instead of numbers:
     col.axis = "white",
     tck = 0.02)

title(ylab = "Technical Difficulty", 
      line = 4, 
      cex.lab = 1)

title(xlab = "Degree of Collaboration", 
      line = 3.5, 
      cex.lab = 1)
###############################################################
# add tick labels
###############################################################
text(x = 1:5,
     y = c(-0.65, rep(-0.7, 4)), 
     cex = 0.7,
     labels = c("Personal", 
                "Lab \nMembers", 
                "Other \nLabs", 
                "EEB \nCommunity", 
                "All \nUsers"))
text(x = c(1, 3, 5),
     y = -1.7, 
     labels = c("Low", "Medium", "High"))
# Technical difficulty levels from https://www.londonschool.com/level-scale/
text(x = 0.2,
     y = seq(1, 13, length.out = 5),
     cex = 0.7,
     adj = 1,
     labels = c("Beginner",
                "Low\nIntermediate",
                "Intermediate", 
                "Pre \nAdvanced",
                "Advanced"))
###############################################################
# plot scatter blobs on sequential cognitive load
###############################################################
for (i in seq(nrow(my_data))) {
  X0 <- my_data$x1[i]
  X1 <- my_data$x2[i]
  Y <- my_data$y_seq[i]
  if (X0 == X1) {
    X0 <- X0 - 0.01
    X1 <- X1 + 0.01
   }
  segments(x0 = X0,
           x1 = X1,
           y0 = Y,
           col = my_colors[i],
           lwd = 23)
}
###############################################################
# add github uses text
###############################################################
text(x = 5.5,
    y = my_data$y_seq,
    labels = my_data$`Ways to use GitHub`,
    adj = 0,
    col = my_colors)
dev.off()
```

![Degree of collaboration *versus* technical difficulty (sequential)](../content/images/scatterblob_1-viridis-turbo.png)

>>>>

---
output:
  bookdown::pdf_document2:
      latex_engine: xelatex
      number_sections: false
      toc: false
mainfont: "Times New Roman"
geometry: margin = 1in
spacing: double
fontsize: 11pt
editor_options:
  markdown:
    wrap: sentence
csl: https://raw.githubusercontent.com/marlonecobos/kuenm/master/ecography.csl
link-citations: yes
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}\doublespacing
   - \usepackage{parskip}
   - \setlength{\parindent}{0em}
   - \setlength{\parskip}{1em}
   - \usepackage{caption}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

@@

We aimed at distributing the use cases discussed in our study across gradients of technical difficulty and collaboration.
For this, we estimated the perceived degree of collaboration and the perceived level of technical difficulty for each use case.

**Degree of collaboration**

_Each author attributed a value ranging from 1 to 10 to their perceived degree of collaboration for a use case, with one representing a use case that offer very low advantages in terms of collaboration, while ten classified that use case as highly advantageous in terms of collaboration._

We determined the degree of collaboration for each GitHub use case as the range of collaborative contexts that can be achieved with each tool, as perceived by the authors. We identified the five levels of possible collaboration in the context of Ecology and Evolutionary Biology research, as follows:

| Level |           Degree of collaboration |
| ----: | --------------------------------: |
|     1 |                        Individual |
|     2 |     Lab or group of collaborators |
|     3 | An organisation, or multiple labs |
|     4 |            The research community |
|     5 |                The general public |

We determined the minimum and maximum degree of collaboration from these levels for each GitHub use case.

**Technical difficulty**

We then assessed the technical difficulty of GitHub use cases in terms of their cognitive load.

As a note, we measured the perceived technical difficulty of each GitHub use case as the degree of difficulty as perceived by the authors, based on personal experience. This metric represents a consensus value among the authors, though the perceived degree of difficulty will ultimately vary among GitHub users depending on each person's previous experience with these use cases.

The table we used to quantify cognitive load was assembled collaboratively using Google Sheets, then manually saved as a `CSV` file called `cognitive_load_table.csv` in the `data` folder.

The collaborative Google Sheet used to quantify cognitive load is also available at the following link: [https://docs.google.com/spreadsheets/d/1YipCCdg5Z8w-IQO67rFfMBSTpLjO-_7HiW44gqzTXPM/edit#gid=0](https://docs.google.com/spreadsheets/d/1YipCCdg5Z8w-IQO67rFfMBSTpLjO-_7HiW44gqzTXPM/edit#gid=0).

\noindent
\newpage

@@

Here, we provide the annotated `R` code and instructions that are required to load and present the indices of technical difficulty and degree of collaboration of the eleven use cases of GitHub we present in our piece.

We used Google Docs and Google Sheets to collect and manipulate data, and plotted results using the basic visualisation functions from `R` version 4.02.

### Data collection

These data were assembled collaboratively using Google Docs following the methodology described in Appendix S1.1. then manually saved as a `CSV` file called `scatterblob_data_raw.csv` in the `data` folder. 
The collaborative Google Spreadsheet is also accessible at the following link: [https://docs.google.com/spreadsheets/d/1zCRgctjfjE2xZN9MkqfFh222K6bvVvA3gT0UdkL_CS8/edit#gid=0](https://docs.google.com/spreadsheets/d/1zCRgctjfjE2xZN9MkqfFh222K6bvVvA3gT0UdkL_CS8/edit#gid=0).

The data table features a first column listing the ways to use GitHub that were addressed in the manuscript (`r paste0(uses[1:13], collapse = "; " )`).

### Creating Figure 2: A summary of ways GitHub can be used showing technical difficulty and degree of collaboration for each use case.

1. Loading the data `CSV` file:

```{r, message=FALSE, collapse=TRUE}
scatterblob_data <- read_csv(here("data", "scatterblob_data_raw.csv"))
```

2. Cleaning the raw data and saving it as a `my_data` object:

```{r}
# names(scatterblob_data)
scatterblob_data$x1 <- scatterblob_data$`Min degree of collaboration X1`
scatterblob_data$x2 <- scatterblob_data$`Max degree of collaboration X2`
scatterblob_data$y_real <- scatterblob_data$`Cognitive load - real - non-overlapping`
scatterblob_data$y_seq <- scatterblob_data$`Cognitive load - sequential`
my_data <- scatterblob_data[1:13,c("Ways to use GitHub","y_seq", "y_real", "x1", "x2")]
```

@@

3. Plotting cognitive load

We used the color palette "turbo" from the R package [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html#the-color-scales):

```{r}
my_colors <- viridis::turbo(n = 100, alpha = 0.9)[c(2, 5, 15, 26, 28, 30,
                           68, 72, 76, 80, 84, 94, 100)]
plot_name <- "scatterblob_1-viridis-turbo"

###############################################################
# start the png connection and set graphical parameters
###############################################################

plot_format <- "png"
if (plot_format == "png") {
  png(file = here(outpath, paste0(plot_name, ".png")),
      width = 8.25,
      height = 6,
      units = "in",
      res = 300,
      bg = "white")  
}
par(xpd = NA,
    mai = c(1.02, 1, 0.82, 4),
    bty="n")

###############################################################
# create the plot background
###############################################################
plot(x = c(0.5, 5.5),
     y = c(0.5, 13.5),
     xlab = "",
     ylab = "",
     col = "white",
     # "hide" the tick labels so we can put some words instead of numbers:
     col.axis = "white",
     tck = 0.02)
title(ylab = "Technical Difficulty", 
      line = 4, 
      cex.lab = 1)
title(xlab = "Degree of Collaboration", 
      line = 3.5, 
      cex.lab = 1)

###############################################################
# add tick labels
###############################################################

text(x = 1:5,
     y = c(-0.65, rep(-0.7, 4)), 
     cex = 0.7,
     labels = c("Personal", 
                "Lab \nMembers", 
                "Other \nLabs", 
                "EEB \nCommunity", 
                "All \nUsers"))
text(x = c(1, 3, 5),
     y = -1.7, 
     labels = c("Low", "Medium", "High"))

# Technical difficulty levels from https://www.londonschool.com/level-scale/
text(x = 0.2,
     y = seq(1, 13, length.out = 5),
     cex = 0.7,
     adj = 1,
     labels = c("Beginner",
                "Low\nIntermediate",
                "Intermediate", 
                "Pre \nAdvanced",
                "Advanced"))

###############################################################
# plot scatter blobs on sequential cognitive load
###############################################################
for (i in seq(nrow(my_data))) {
  X0 <- my_data$x1[i]
  X1 <- my_data$x2[i]
  Y <- my_data$y_seq[i]
  if (X0 == X1) {
    X0 <- X0 - 0.01
    X1 <- X1 + 0.01
   }
  segments(x0 = X0,
           x1 = X1,
           y0 = Y,
           col = my_colors[i],
           lwd = 23)
}

###############################################################
# add github uses text
###############################################################
text(x = 5.5,
    y = my_data$y_seq,
    labels = my_data$`Ways to use GitHub`,
    adj = 0,
    col = my_colors)
dev.off()
```

![A summary of ways GitHub can be used showing technical difficulty and degree of collaboration for each use case.](../content/images/scatterblob_1-viridis-turbo.png)
